## Customizability

`makeParallel` allows us to pass in functions that customize the behavior.
For example, we may want the generated code to explicitly set the number of parallel workers to `N` as specified by the option `mc.cores` in the parallel package.
The following function prepends the expression `options(mc.cores = N)` to the generated code.

```{r}
coresGenerate <- function(schedule, mc.cores = 2L, ...)
{
    # Rely on the method dispatch for the actual work.
    out <- generate(schedule, ...)

    # Construct an expression containing the desired code.
    setCores <- substitute(options(mc.cores = MC_CORES)
                          , list(MC_CORES = mc.cores))

    # Combine the newly constructed expression with what would have been
    # generated otherwise.
    out@code <- c(setCores, writeCode(out))
    out
}
```

We can use this function as follows:

```{r}
lapplyCode <- parse(text = "
    x <- list(a = 1:10, beta = exp(-3:3), logic = c(TRUE,FALSE,FALSE,TRUE))
    m1 <- lapply(x, mean)
")

transformed <- makeParallel(lapplyCode, generator = coresGenerate,
                            generatorArgs = list(mc.cores = 3L))
```

When we extract the code from this object with `writeCode` we see that it sets `options(mc.cores = 3L)` as the first line.

```{r}
writeCode(transformed)
```

```{r, echo = FALSE, results = "hide"}
# Testing, make sure the docs do what they say!
stopifnot(writeCode(transformed)[[1]] == quote(options(mc.cores = 3L)))
```


## Extensibility

Some schedulers must be tied to their code generators.
`inferGraph, schedule`, and `generate` are all generic functions, so we can allow user defined classes to extend the system through R's S4 object oriented programming system.

Building on the example above, we can define a class `WorkerMapSchedule` containing `MapSchedule` that adds a slot for `mc.cores`.

```{r}
setClass("WorkerMapSchedule", slots = c(mc.cores = "integer"), contains = "MapSchedule")
```

Here's a helper constructor function:

```{r}
workerMapSchedule = function(graph, mc.cores = 2L, ...)
{
    message(sprintf("User defined scheduler, mc.cores = %s", mc.cores))
    out = mapSchedule(graph, ...)
    new("WorkerMapSchedule", out, mc.cores = mc.cores)
}
```

Now we need to associate a code generator with `WorkerMapSchedule`.
We can use an existing one or we can define our own.
Since we've already defined `coresGenerate` above we'll just wrap that.

```{r}
setMethod("generate", "WorkerMapSchedule", function(schedule, ...)
    coresGenerate(as(schedule, "MapSchedule"), mc.cores = schedule@mc.cores, ...)
)
```

Finally, we can use the code as follows:

```{r}
transformed <- makeParallel(code, scheduler = workerMapSchedule, mc.cores = 3L)

writeCode(transformed)
```

```{r, echo = FALSE, results = "hide"}
# Testing, make sure the docs do what they say!
stopifnot(writeCode(transformed)[[1]] == quote(options(mc.cores = 3L)))
```


In this section we went beyond the basic customization in the previous section in two ways.
First, we extended the existing class hierarchy by defining our own scheduler.
Second, we defined methods and relied on method dispatch to control some aspects of the code generation process.
We did not have to touch the dependency graph computations.


