---
title: "makeParallel quick start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{makeParallel-quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

> This vignette provides a brief introduction to using the makeParallel
> package. If you would like to learn the details then see the vignette
> titled "makeParallel concepts", `vignette("makeParallel-concepts")`.

## background

If you're totally new to parallel programming then read this first.

Nearly all modern desktop and laptop computers have multiple physical
processors. Parallel code uses multiple physical processors to do several
computations at the same time. This is sometimes faster, but not always. R
usually requires programmers to explicitly use parallelism, for example
with the recommended `parallel` package. Writing parallel code requires
more expertise from us, the users. The purpose of this package is to
generate more efficient parallel code so that we don't have to.

## quick start

Let's start with an existing R script `mp_example.R` in an otherwise empty
directory.

```{r}
d <- tempdir()
oldscript <- system.file("tests/testthat/mp_example.R", package = "makeParallel")
script <- file.path(d, "mp_example.R")
file.copy(oldscript, script)
list.files(d)
```

The function `makeParallel` wraps most of the other steps and forms the
core of this package. As the name suggests, this function makes parallel
code from serial R code. We can use it on an existing script as follows:

```{r}
library(makeParallel)

g <- makeParallel(script, file = TRUE)
```

Specifying `file = TRUE` means that `makeParallel` will write the newly
generated code into a file based on the original name.

```{r}
list.files(d)
```

```{r, echo = FALSE, results = "hide"}
stopifnot("gen_mp_example.R" %in% list.files(d))
```

<!--
`makeParallel` returns an object of class `GeneratedCode` which contains
the newly created executable code as well as the results of the
intermediate steps. 
-->

`makeParallel` can directly replace serial code with parallel versions. The
following example comes from the documentation for `lapply`:

```{r}
input <- parse(text = "
    x <- list(a = 1:10, beta = exp(-3:3), logic = c(TRUE,FALSE,FALSE,TRUE))
    m1 <- lapply(x, mean)
    m2 <- list()
    for(i in seq_along(x)) {
        m2[[i]] = mean(x[[i]])
    }
    ")

transformed <- makeParallel(input)
```

The resulting object has transformed both the `for` loop and the `lapply`
to `parallel::mclapply`. Note this will not run in parallel on a Windows
system. A future release will support Windows. 

Here is the code we started with.

```{r}
input
```

`makeParallel` then generated this new code.

```{r}
newcode <- writeCode(transformed)

newcode
```

We can evaluate the new code and inspect the object `m2`.

```{r}
eval(newcode)
m2
```

We can also evaluate the original code and verify that it produces the same
output.

```{r}
m2new <- m2
eval(input)
m2

stopifnot( all(as.numeric(m2) == as.numeric(m2new)) )
```

Ready to learn more? Read on in the "makeParallel concepts" vignette,
`vignette("makeParallel-concepts")`.

```{r, echo = FALSE, results = "hide"}
setwd(olddir)
unlink(td, recursive = TRUE)
```
